---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Some preparations
```{r, setup, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
Sys.setlocale("LC_TIME", "C")
```

## Loading and preprocessing the data
### Load data
```{r, load, echo=TRUE}
steps <- tbl_df(
    read.csv(
        unz("activity.zip", "activity.csv"),
        colClasses = c("integer", "factor", "factor")
    )
) 
```

### Data structure
```{r, structure, echo=TRUE}
str(steps)
```


## What is mean total number of steps taken per day?
### The total number of steps taken per day
```{r, steps_per_date, echo=TRUE}
steps_per_date <- 
    steps %>%
        select(date, steps) %>%
        group_by(date) %>%
        summarize(total_steps = sum(steps, na.rm = TRUE)) %>%
        mutate(date = as.Date(date))

head(steps_per_date)
```

### Histogram of the total number of steps taken each day
#### Histogram
```{r, histogram, echo=TRUE}
with(
    steps_per_date,
    hist(
        total_steps,
        xlab = "Total steps",
        main = "Histogram of the total number of steps taken each day"
    )
)
```

#### Barplot
```{r, barplot, echo=TRUE}
with(
    steps_per_date,
    barplot(
        total_steps, names.arg = format(date, "%b %d"),
        xlab = "Date", ylab = "Total steps",
        main = "Histogram of the total number of steps taken each day"
    )
)
```

#### Result histogram
```{r, plot, echo=TRUE}
with(
    steps_per_date,
    plot(
        date, total_steps, type = "h",
        xlab = "Date", ylab = "Total steps",
        main = "Histogram of the total number of steps taken each day"
    )
)
```


### Mean of the total number of steps taken per day
```{r, mean_steps_per_date, echo=TRUE}
mean_orig <- 
    with(
        steps_per_date,
        mean(total_steps, na.rm = TRUE)
    )

print(mean_orig)
```

### Median of the total number of steps taken per day
```{r, meadian_steps_per_date, echo=TRUE}
median_orig <- 
    with(
        steps_per_date,
        median(total_steps, na.rm = TRUE)
    )

print(median_orig)
```

## What is the average daily activity pattern?
### The average number of steps taken, averaged across all days
```{r, steps_by_interval, echo=TRUE}
steps_by_interval <- 
    steps %>%
        select(interval, steps) %>%
        group_by(interval) %>%
        summarize(avg_steps = mean(steps, na.rm = TRUE))

head(steps_by_interval)
```
   
### Plot
```{r, steps_per_date_plot, echo=TRUE}
with(
    steps_by_interval,
    plot(
        as.integer(interval), avg_steps, type = 'l', col = interval,
        main = "Daily activity pattern", 
        xlab = "Interval", ylab = "Number of steps"
    )
)
```

### 5-minute interval which contains the maximum number of steps
```{r, five_minute_interval, echo=TRUE}
steps_by_interval %>%
    filter(avg_steps == max(avg_steps)) %>%
    print
```

## Imputing missing values
### The total number of missing values in the dataset
```{r, total_missing_values, echo=TRUE}
steps %>%
    filter(is.na(steps)) %>%
    count
```


### Filling in all of the missing values in the dataset
```{r, filling_missing_values, echo=TRUE}
aveges <- steps_by_interval$avg_steps
names(aveges) <- steps_by_interval$interval

new_steps <-
    steps %>%
        mutate(
            steps = ifelse(
                is.na(steps),
                aveges[interval],
                steps
            )
        )

head(new_steps)
```

### The total number of steps taken each day
```{r, total_number_of_steps_on_filled_data}
new_steps_per_date <- 
    new_steps %>%
        select(date, steps) %>%
        group_by(date) %>%
        summarize(total_steps = sum(steps, na.rm = TRUE)) %>%
        mutate(date = as.Date(date))

head(new_steps_per_date)
```

### Histogram of the total number of steps taken per day
```{r, new_plot, echo=TRUE}
with(
    new_steps_per_date,
    plot(
        date, total_steps, type = "h",
        xlab = "Date", ylab = "Total steps",
        main = "Histogram of the total number of steps taken each day"
    )
)
```

### Mean of the total number of steps taken per day
```{r, new_mean_steps_per_date, echo=TRUE}
mean_new <-
    with(
        new_steps_per_date,
        mean(total_steps)
    )

print(mean_new)
```

### Median of the total number of steps taken per day
```{r, new_meadian_steps_per_date, echo=TRUE}
median_new <-
    with(
        new_steps_per_date,
        median(total_steps)
    )

print(median_new)
```

### Conclusions
1. The mean and median of the total number of steps taken per day on filled data list differs from the mean and median on original data.
```{r, conclusions_1, echo=TRUE}
abs(median_new - median_orig)

abs(mean_new - mean_orig)
```

2. The mean and median of the total number of steps taken per day on filled data list are equals:
```{r, conclusions_2, echo=TRUE}
median_new - mean_new
```


## Are there differences in activity patterns between weekdays and weekends?
### Mutate data
```{r, actitvity_by_daytype, echo=TRUE}
actitvity_by_daytype <-
    new_steps %>%
        mutate(
            daytype = as.factor(
                ifelse(
                    weekdays(
                        as.Date(date),
                        abbreviate = TRUE
                    ) %in% c("Sat", "Sun"),
                    "weekend",
                    "weekday"
                )
            )
        )

str(actitvity_by_daytype)

actitvity_by_daytype_groupped <-
    actitvity_by_daytype %>%
        group_by(daytype, interval) %>%
        summarize(avg_steps = mean(steps))

head(actitvity_by_daytype_groupped)
```

### Plot
```{r, actitvity_by_daytype_plot, echo=TRUE}
par(mfrow = c(2, 1))

with(
    actitvity_by_daytype_groupped %>% filter(daytype == "weekend"),
    plot(
        as.integer(interval), avg_steps, type = 'l',
        main = "Weekends", 
        xlab = "Interval", ylab = "Number of steps"
    )
)

with(
    actitvity_by_daytype_groupped %>% filter(daytype == "weekday"),
    plot(
        as.integer(interval), avg_steps, type = 'l',
        main = "Weekdays", 
        xlab = "Interval", ylab = "Number of steps"
    )
)
```

### Plot 2
```{r, actitvity_by_daytype_plot_2, echo=TRUE}
library(ggplot2)
ggplot2::ggplot(
    actitvity_by_daytype_groupped,
    ggplot2::aes(x = as.integer(interval), y = avg_steps)
) + 
    ggplot2::facet_grid(daytype ~ .) +
    ggplot2::xlab("Interval") + 
    ggplot2::ylab("Number of steps") + 
    ggplot2::geom_line()
```
